---
title: "Analyzing Taylor Swift's Midnights Album"
author:
  - name: "Marc Shervin Ignacio"
    affiliation: "2019-04690"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    theme: lumen
    toc: yes
    toc_float: yes
    toc_depth: 4
    number_sections: no
    highlight: pygments
    fig_caption: yes
editor_options: 
  markdown: 
    wrap: 72
---

### What is Topic Modeling?

<p>Topic modeling is a machine learning technique that automatically
analyzes text data to determine cluster words for a set of documents.
This is known as 'unsupervised' machine learning because it doesn't
require a predefined list of tags or training data that's been
previously classified by humans.</p>

<div>

For this project, we are going to use Latent Dirichlet Allocation (LDA).
Simply:

Taken from:

[A Beginner's Guide to Latent Dirichlet Allocation(LDA) \| by Ria
Kulshrestha \| Towards Data
Science](https://towardsdatascience.com/latent-dirichlet-allocation-lda-9d1cd064ffa2)

# **LDA**

It is one of the most popular topic modeling methods. Each document is
made up of various words, and each topic also has various words
belonging to it. The aim of LDA is to find topics a document belongs to,
based on the words in it.

</div>

Onto the project:\

```{r setup}
require('topicmodels')
require(knitr)
require(kableExtra)
require(geniusr)
library(topicmodels)
library(tidyr)
library(tidytext)
library(dplyr)
library(ggplot2)
```

```{r}
Midnights_album <- get_album_tracklist_search(artist_name = "Taylor Swift", album_name = "Midnights")
```

```{r}
lyrics <- rbind(get_lyrics_url(Midnights_album$song_lyrics_url[1])[c(1,4)],
           get_lyrics_url(Midnights_album$song_lyrics_url[2])[c(1,4)],
               get_lyrics_url(Midnights_album$song_lyrics_url[3])[c(1,4)],
               get_lyrics_url(Midnights_album$song_lyrics_url[4])[c(1,4)],
               get_lyrics_url(Midnights_album$song_lyrics_url[5])[c(1,4)],
               get_lyrics_url(Midnights_album$song_lyrics_url[6])[c(1,4)],
               get_lyrics_url(Midnights_album$song_lyrics_url[7])[c(1,4)],
               get_lyrics_url(Midnights_album$song_lyrics_url[8])[c(1,4)],
               get_lyrics_url(Midnights_album$song_lyrics_url[9])[c(1,4)],
               get_lyrics_url(Midnights_album$song_lyrics_url[11])[c(1,4)],
               get_lyrics_url(Midnights_album$song_lyrics_url[12])[c(1,4)],
            	get_lyrics_url(Midnights_album$song_lyrics_url[13])[c(1,4)])

```

```{r}
colnames(lyrics) <- c("text","id")
```

```{r}
head(lyrics, 20) %>%
  kbl(caption = "Midnights album") %>%
  kable_classic(full_width = F, html_font = "Cambria")
```

```{r}
# function to expand contractions in an English-language source
fix.contractions <- function(doc) {
  # "won't" is a special case as it does not expand to "wo not"
  doc <- gsub("won't", "will not", doc)
  doc <- gsub("can't", "can not", doc)
  doc <- gsub("n't", " not", doc)
  doc <- gsub("'ll", " will", doc)
  doc <- gsub("'re", " are", doc)
  doc <- gsub("'ve", " have", doc)
  doc <- gsub("'m", " am", doc)
  doc <- gsub("'d", " would", doc)
  # 's could be 'is' or could be possessive: it has no expansion
  doc <- gsub("'s", " is", doc)
  return(doc)
}
dup_lyrics <- lyrics
dup_lyrics$text <- sapply(dup_lyrics$text,fix.contractions)

filtered_lyrics <- dup_lyrics %>% 
  unnest_tokens(input=text, output=word) %>% #Prepare the data
  anti_join(stop_words)#Remove the stop words

my_colors <- c("#E69F00", "#56B4E9", "#009E73", "#CC79A7", "#D55E00")
filtered_lyrics %>%
  count(word, sort = TRUE) %>%
  top_n(10) %>%
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot() +
    geom_col(aes(word, n), fill = my_colors[4]) +
    theme(legend.position = "none", 
          plot.title = element_text(hjust = 0.5),
          panel.grid.major = element_blank()) +
    xlab("") + 
    ylab("Song Count") +
    ggtitle("Most Frequently Used Words in Taylor Swift Lyrics") +
    coord_flip()


```

### Analysis

```{r}
dtm <- dup_lyrics %>% 
  unnest_tokens(input=text, output=word) %>% #Prepare the data
  anti_join(stop_words) %>% #Remove the stop words
  count(id, word) %>%  #Count the instances of our words
  cast_dtm(document=id, term=word, value=n) #Cast into a doc-term matrix


```

Here we create our models:

```{r}
mod2 <- LDA(x=dtm, k=2, method="Gibbs", 
            control=list(alpha=1))
mod3 <- LDA(x=dtm, k=3, method="Gibbs", 
            control=list(alpha=1))
mod4 <- LDA(x=dtm, k=4, method="Gibbs", 
            control=list(alpha=1))
mod5 <- LDA(x=dtm, k=5, method="Gibbs", 
            control=list(alpha=1))
mod6 <- LDA(x=dtm, k=6, method="Gibbs", 
            control=list(alpha=1))
mod7 <- LDA(x=dtm, k=7, method="Gibbs", 
            control=list(alpha=1))
mod8 <- LDA(x=dtm, k=8, method="Gibbs", 
            control=list(alpha=1))
mod9 <- LDA(x=dtm, k=9, method="Gibbs", 
            control=list(alpha=1))
mod10 <- LDA(x=dtm, k=10, method="Gibbs", 
           control=list(alpha=1))
mod11 <- LDA(x=dtm, k=11, method="Gibbs", 
           control=list(alpha=1))
mod12 <- LDA(x=dtm, k=12, method="Gibbs", 
           control=list(alpha=1))
mod13 <- LDA(x=dtm, k=13, method="Gibbs", 
           control=list(alpha=1))
mod14 <- LDA(x=dtm, k=14, method="Gibbs", 
           control=list(alpha=1))
mod15 <- LDA(x=dtm, k=15, method="Gibbs", 
           control=list(alpha=1))
mod16 <- LDA(x=dtm, k=16, method="Gibbs", 
           control=list(alpha=1))
mod17 <- LDA(x=dtm, k=17, method="Gibbs", 
           control=list(alpha=1))

models <- c(mod2, mod3, mod4, mod5, mod6, mod7, mod8, mod9, mod10, mod11, mod12, mod13, mod14, mod15, mod16, mod17)
```

```{r}
#Here we create a list of our models and their log_likelihood and perplexity scores
container <- list()
for(i in models){
  container <- append(container,
                      list(list(k=i@k, model=i, log_likelihood=topicmodels::logLik(i), 
                           perplexity = perplexity(object=i, newdata=dtm)))
                      )
}
```

```{r}
#Plotting our perplexity score
x <- sapply(container, '[[', 'k')
y <- sapply(container, '[[', 'perplexity')
plot(x, y, xlab="number of clusters, k", ylab="perplexity score", type="o")
namebank = as.character(2:17)
text(x, y, labels=namebank, cex= 0.7, pos=3)
```

Our desired \# of topics is 11:

```{r}
kable(terms(mod11, k=10))
```

```{r}
new_perplexity_score <- numeric(length(container))

# Run each model for 100 iterations
for (i in seq_along(container)) {
  mod2 <- LDA(x=dtm, model=container[[i]]$model,
              control=list(iter=5000, thin=1))
  new_perplexity_score[i] <- perplexity(object=mod2, newdata=dtm)
}

k <- 2:17
plot(x=k, y=new_perplexity_score, xlab="number of clusters, k", 
     ylab="perplexity score", type="o")
namebank = as.character(2:17)
text(x, y, labels=namebank, cex= 0.7, pos=3)
```

 

```{r}
#Wordclouds for Model with 5 topics
library(wordcloud)
# Display wordclouds one at a time
for (j in 1:11) {
  # Generate a table with word frequences for topic j
  word_frequencies <- tidy(mod11, matrix="beta") %>% 
    mutate(n = trunc(beta * 10000)) %>% 
    filter(topic == j)

  # Display word cloud
  layout(matrix(c(1, 2), nrow=2), heights=c(1, 4))
  par(mar=rep(0, 4))
  plot.new()
  text(x=0.5, y=0.5, paste("Topic ",j))
  wordcloud(words = word_frequencies$term, 
            freq = word_frequencies$n,
            max.words = 20,
            scale = c(3, 0.5),
            colors = c("DarkOrange", "CornflowerBlue", "DarkRed"), 
            rot.per = 0.3)
}
```
